Forcing for Isabelle

### 2018-12-21 LICS/recursion.tex
Fixed typo.
Hablando de interfaz (por hacer)

We close our discussion of names and their values by making explicit
the names for elements in $M$, once more we refer to
\cite{2018arXiv180705174G} for our formalization. The definition of
$\chk(x)$ is a straightforward $\in$-recursion:
\begin{equation*}
  \label{eq:def-check}
  \chk(x)\defi\{\lb\chk(y),\1\rb : y\in x\}
\end{equation*}
It can be easily proved by $\in$-induction that $\val(G,\chk(x))=x$.
But to conclude $M\subseteq M[G]$ one also needs to have
$\chk(x) \in M$; this result requires the internalization of
recursively defined functions. This is also needd to prove
$G \in M[G]$; let us define
$\dot G= \{\lb \chk(p),p\rb : p \in \PP \}$, it is easy to prove
$\val(G,\dot G) = G$. Proving $\dot G\in M$ involves
knowing $\chk(x) \in M$ and using one instance of Replacement.

Paulson proved absoluteness results for definitions by recursion and
one of our next goals is to instantiate the appropiate locale
\isa{M{\isacharunderscore}eclose} which is the last layer of a pile of
locales; it will take us some time to prove that any ctm satisfies the
assumptions involved in those locales. In the meantime, we declare a
locale \isa{M{\isacharunderscore}extra{\isacharunderscore}assms}
assembling both assumptions.

>>>>

We close our discussion of names and their values by making explicit
the names for elements in $M$, once more we refer to
\cite{2018arXiv180705174G} for our formalization. The definition of
$\chk(x)$ is a straightforward $\in$-recursion:
\begin{equation*}
  \label{eq:def-check}
  \chk(x)\defi\{\lb\chk(y),\1\rb : y\in x\}
\end{equation*}
It can be easily proved by $\in$-induction that $\val(G,\chk(x))=x$.
But to conclude $M\subseteq M[G]$ one also needs to have
$\chk(x) \in M$; this result requires the internalization of
recursively defined functions. This is also needed to prove
$G \in M[G]$; let us define
$\dot G= \{\lb \chk(p),p\rb : p \in \PP \}$, it is easy to prove
$\val(G,\dot G) = G$. Proving $\dot G\in M$ involves knowing
$\chk(x) \in M$ and using one instance of Replacement.

Paulson proved absoluteness results for definitions by recursion and
one of our next goals is to instantiate the appropiate locale
\isa{M{\isacharunderscore}eclose} which is the last layer of a pile of
locales. It will take us some time to prove that any ctm satisfies the
assumptions involved in those locales; as we mentioned, Paulson work
is mostly done at the meta-level, i.e. the assumptions are instances
of Separation and Replacement where the predicates and functions are
Isabelle functions of type \isa{i{\isasymRightarrow}i} and
\isa{i{\isasymRightarrow}o}, respectively. In contrast, we assume that
$M$ is a model of $ZF$, therefore to deduce that $M$ satisfies a
Separation instance, we have to define an internalized formula whose
satisfaction is equivalent to the meta-level predicate (cf. the
interface described in Section~\ref{sec:axioms-models-set-theory} and
also the concrete example given in the proof of Union).

In the meantime, we declare a locale
\isa{M{\isacharunderscore}extra{\isacharunderscore}assms} assembling
both assumptions ($M$ closed under $\chk$ and the instance of
Replacement); in this paper we explicitly mention where we use them.
